---
title: "NMA Project"
author: "Omar Jouini"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```

installing the needed libraries
```{r}
#install.packages("multinma")
```

loading the needed libraries and setting the CPU Cores to a parrallel usage
```{r}
library(multinma)
library(tidyverse)
library(janitor)

options(mc.cores = parallel::detectCores())
```

# Loading the data

Loading the parkinsons data. The data are available in this package as `parkinsons`

```{r}
head(parkinsons)

# infos about the dataset
parkinsons
```

# starting the analysis

## Arm based data
means of y and the Standard error

Preparations:
In the dataset we have continous data of the "off-time" reducion. 
"off-time" referes to the time where parkinsonian drugs are not in effect, to which the patients experiances parkinsonian effects. The reduction of the off time is a therapeutic goal.

We use the function `set_agd_arm()` to set up the network.

```{r}
arm_net <- set_agd_arm( data = parkinsons,
                        study = studyn,
                        trt = trtn,
                        y = y,
                        se= se,
                        sample_size = n)
arm_net
```
networks can be ploted
```{r Plot}
plot(arm_net)
plot(arm_net, weight_edges = T)
plot(arm_net, weight_nodes = T)
plot(arm_net, weight_edges = T, weight_nodes = T)
```
### Meta-Analisis Models
Both fixed effect (FE) and randome effects (RE) Models are possible,  but fixed effects Models are missing.

#### FE-Model
```{r}
summary(normal(scale = 100))


arm_fit_FE <- nma(arm_net, 
                  trt_effects = "fixed",
                  prior_intercept = normal(scale = 100),
                  prior_trt = normal(scale = 10))

arm_fit_FE
#By default, summaries of the study-specific intercepts μ  are hidden, but could be examined by changing the pars argument:

print(arm_fit_FE, pars = c("d", "mu"))

plot_prior_posterior(arm_fit_FE)

```

### RE Model

Here we also set a prior for the heterogenity

```{r RE Model}

arm_fit_RE <- nma(arm_net, 
                  seed = 379394727,
                  trt_effects = "random",
                  prior_intercept = normal(scale = 100),
                  prior_trt = normal(scale = 100),
                  prior_het = half_normal(scale = 5),
                  adapt_delta = 0.99)



pairs(arm_fit_RE, pars = c("mu[4]", "d[3]", "delta[4: 3]", "tau"))
```

### Model comparison
 it is possible to do a comparison withth dic() "Deviance Information Criterion (DIC)" function:
 
```{r}
arm_dic_FE <- dic(arm_fit_FE)
arm_dic_FE

arm_dic_RE <- dic(arm_fit_RE)
arm_dic_RE

plot(arm_dic_FE)
plot(arm_dic_RE)

```
### Further results 
The package also alows you to produce further results, here erlative effects against a reference, where you can define the reference

```{r}
arm_rel_eff_FR <- relative_effects(arm_fit_FE, trt_ref = 1)
arm_rel_eff_FR
plot(arm_rel_eff_FR, ref_line = 0)

arm_releff_RE <- relative_effects(arm_fit_RE, trt_ref = 1)
plot(arm_releff_RE, ref_line = 0)
```


```{r}
arm_pred_FE <- predict(arm_fit_FE, 
                       baseline = distr(qnorm, mean = -0.73, sd = 21^-0.5),
                       type = "response",
                       baseline_trt = 1)
arm_pred_FE
plot(arm_pred_FE)
```

### The treatments can also be ranked

```{r}
arm_ranks <- posterior_ranks(arm_fit_FE) #tretmant ranking : produces posterior rankings, which have a distribution (e.g. mean/median rank and 95% Credible Interval)
arm_rankprobs <- posterior_rank_probs(arm_fit_FE) # probabilities, which give the posterior probabilities of being ranked first, second, etc. out of all treatments.
plot(arm_rankprobs)
plot(arm_ranks)
# also in a cummulativ way
arm_cumrankprobs <- posterior_rank_probs(arm_fit_FE, cumulative = TRUE)
plot(arm_cumrankprobs)
```

## Useing contraste based data 
mean difference and standard error difference
```{r}
contr_net <- set_agd_contrast(parkinsons, 
                              study = studyn,
                              trt = trtn,
                              y = diff, 
                              se = se_diff,
                              sample_size = n)
contr_net
plot(contr_net, weight_edges = TRUE, weight_nodes = TRUE)


```
Fixed effects Models

```{r}
contr_fit_FE <- nma(contr_net, 
                    trt_effects = "fixed",
                    prior_trt = normal(scale = 100))
contr_fit_FE

plot_prior_posterior(contr_fit_FE)
```
random effects Models

```{r}
contr_fit_RE <- nma(contr_net, 
                    seed = 1150676438,
                    trt_effects = "random",
                    prior_trt = normal(scale = 100),
                    prior_het = half_normal(scale = 5),
                    adapt_delta = 0.99)

pairs(contr_fit_RE, pars = c("d[3]", "delta[4: 4 vs. 3]", "tau"))

contr_fit_RE
```

### Model comparison
```{r}
contr_dic_FE <- dic(contr_fit_FE)
contr_dic_RE <- dic(contr_fit_RE)
plot(contr_dic_FE)
plot(contr_dic_RE)
```

### Further Result

```{r}
contr_releff_FE <- relative_effects(contr_fit_FE, trt_ref = 1)

contr_releff_RE <- relative_effects(contr_fit_RE, trt_rpef = 1)
plot(contr_releff_FE, ref_line = 0)
plot(contr_releff_RE, ref_line = 0)
```
#### Prediction

```{r}
contr_pred_FE <- predict(contr_fit_FE, 
                       baseline = distr(qnorm, mean = -0.73, sd = 21^-0.5),
                       type = "response",
                       baseline_trt = 1)

contr_pred_RE <- predict(contr_fit_RE, 
                       baseline = distr(qnorm, mean = -0.73, sd = 21^-0.5),
                       type = "response",
                       baseline_trt = 1)
contr_pred_FE
plot(contr_pred_FE)

contr_pred_RE
plot(contr_pred_RE)
```


## Combination of both

```{r}
studies <- parkinsons$studyn
(parkinsons_arm <- parkinsons[studies %in% 1:3, ])
(parkinsons_contr <- parkinsons[studies %in% 4:7, ])
```


```{r}
mix_arm_net <- set_agd_arm(parkinsons_arm, 
                           study = studyn,
                           trt = trtn,
                           y = y, 
                           se = se,
                           sample_size = n)

mix_contr_net <- set_agd_contrast(parkinsons_contr, 
                                  study = studyn,
                                  trt = trtn,
                                  y = diff, 
                                  se = se_diff,
                                  sample_size = n)

mix_net <- combine_network(mix_arm_net, mix_contr_net)
mix_net
```
```{r}
mix_fit_FE <- nma(mix_net, 
                  trt_effects = "fixed",
                  prior_intercept = normal(scale = 100),
                  prior_trt = normal(scale = 100))

mix_fit_FE
print(mix_fit_FE, pars = c("d", "mu"))
plot_prior_posterior(mix_fit_FE)

mix_fit_RE <- nma(mix_net, 
                  seed = 437219664,
                  trt_effects = "random",
                  prior_intercept = normal(scale = 100),
                  prior_trt = normal(scale = 100),
                  prior_het = half_normal(scale = 5),
                  adapt_delta = 0.99)

pairs(mix_fit_RE, pars = c("d[3]", "delta[4: 4 vs. 3]", "tau"))
```

### Creating a prior/posterior plot

```{r}
plot_prior_posterior(mix_fit_RE)

```

### model comparrison

```{r}
mix_dic_FE <- dic(mix_fit_FE)
mix_dic_RE <- dic(mix_fit_RE)

plot(mix_dic_FE)
plot(mix_dic_RE)
```

### Further results
#### reative effect
```{r}
mix_releff_FE <- relative_effects(mix_fit_FE, trt_ref = 1)
mix_releff_RE <- relative_effects(mix_fit_RE, trt_ref = 1)
plot(mix_releff_RE, ref_line = 0)
plot(mix_releff_FE, ref_line = 0)
```


#### Predictions

```{r}
mix_pred_FE <- predict(mix_fit_FE, 
                       baseline = distr(qnorm, mean = -0.73, sd = 21^-0.5),
                       type = "response",
                       baseline_trt = 1)
mix_pred_FE

mix_pred_RE <- predict(mix_fit_RE, 
                       baseline = distr(qnorm, mean = -0.73, sd = 21^-0.5),
                       type = "response",
                       baseline_trt = 1)
mix_pred_RE
plot(mix_pred_RE)
plot(mix_pred_FE)
```


# Reproduction with additional Data

```{r addting to the parkinsons dataset}

parkinsons 



```

